Nonlinear mixed effects models for time series forecasting of smart meter demand

### 2020-10-21 paper.Rmd
Minor fixes
- Add extra sentence explaining mixed effects models.
- Correct author details.
- Add github repo location.

---
title: "Nonlinear mixed effects models for time series forecasting of smart meter demand"
author:
- familyname: Roach
  othernames: Cameron
  address: Monash University
  email: cameron.roach@monash.edu
  correspondingauthor: true
- familyname: Hyndman
  othernames: Rob
  address: Monash University
- familyname: Taieb
  othernames: Souhaib Ben
  address: University of Mons
abstract: Buildings are typically equipped with smart meters to measure electricity demand at regular intervals. Smart meter data for a single building have many uses, such as forecasting and assessing overall building performance. However, when data are available from multiple buildings, there are additional applications that are rarely explored. For instance, we can explore how different building characteristics influence energy demand. If each building is treated as a random effect and building characteristics are handled as fixed effects, a mixed effects model can be used to estimate how characteristics affect energy usage. In this paper we demonstrate that producing one day ahead demand predictions for 123 commercial office buildings using mixed models can improve forecasting accuracy. We experiment with random intercept, random intercept and slope, and nonlinear mixed models. The predictive performance of the mixed effects models are tested against naive, linear and nonlinear benchmark models fitted to each building separately. This research justifies using mixed models to improve forecasting accuracy and to quantify changes in energy consumption under different building configuration scenarios.
keywords: "time series forecasting, mixed-effects models, smart meters, energy, electricity"
wpnumber: 1/19
jelcodes: C10,C14,C52
blind: false
cover: false
toc: false
bibliography: ["library.bib", "packages.bib"]
biblio-style: authoryear-comp
output:
  MonashEBSTemplates::workingpaper:
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    includes:
      in_header: preamble.tex
    keep_tex: yes
    number_sections: yes
    citation_package: biblatex
---

@@

Several papers have examined forecasting electricity demand for buildings by fitting separate models to each building [@Ghofrani2011-tb; @Gajowniczek2014-ek; @Arora2016-zh; @Ben_Taieb2016-wl]. While some have attempted to improve forecasts by leveraging the hierarchical nature of electricity demand [@Ben_Taieb2020-it; @Ben_Taieb2017-ok] few, if any, have explored improving forecast accuracy using a mixed effects framework. If buildings behave in a similar manner a well-specified mixed model may produce more accurate forecasts than individual models. Furthermore, a mixed-model framework allows us to quantify differences between buildings which would not otherwise be possible when using a "building-specific" modelling approach. A mixed effects approach opens the door to scenario analyses by allowing us to estimate how demand might change under different equipment or usage scenarios.

@@

Several linear and mixed effects models were tested to determine which produced the most accurate forecasts conditional on selected features. Here we describe the various benchmarking and mixed effects models and their formulations. All analysis was produced using the R statistical programming language [@R-base]. Mixed effects models were fit using the `lme4` package for mixed effects models [@R-lme4].

>>>>

---
title: "Nonlinear mixed effects models for time series forecasting of smart meter demand"
author:
- familyname: Roach
  othernames: Cameron
  address: Monash University
  email: camroach87@gmail.com
  correspondingauthor: true
- familyname: Hyndman
  othernames: Rob
  address: Monash University
- familyname: Ben Taieb
  othernames: Souhaib
  address: University of Mons
abstract: Buildings are typically equipped with smart meters to measure electricity demand at regular intervals. Smart meter data for a single building have many uses, such as forecasting and assessing overall building performance. However, when data are available from multiple buildings, there are additional applications that are rarely explored. For instance, we can explore how different building characteristics influence energy demand. If each building is treated as a random effect and building characteristics are handled as fixed effects, a mixed effects model can be used to estimate how characteristics affect energy usage. In this paper we demonstrate that producing one day ahead demand predictions for 123 commercial office buildings using mixed models can improve forecasting accuracy. We experiment with random intercept, random intercept and slope, and nonlinear mixed models. The predictive performance of the mixed effects models are tested against naive, linear and nonlinear benchmark models fitted to each building separately. This research justifies using mixed models to improve forecasting accuracy and to quantify changes in energy consumption under different building configuration scenarios.
keywords: "time series forecasting, mixed-effects models, smart meters, energy, electricity"
wpnumber: 1/19
jelcodes: C10,C14,C52
blind: false
cover: false
toc: false
bibliography: ["library.bib", "packages.bib"]
biblio-style: authoryear-comp
output:
  MonashEBSTemplates::workingpaper:
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    includes:
      in_header: preamble.tex
    keep_tex: yes
    number_sections: yes
    citation_package: biblatex
---

@@

Several papers have examined forecasting electricity demand for buildings by fitting separate models to each building [@Ghofrani2011-tb; @Gajowniczek2014-ek; @Arora2016-zh; @Ben_Taieb2016-wl]. While some have attempted to improve forecasts by leveraging the hierarchical nature of electricity demand [@Ben_Taieb2020-it; @Ben_Taieb2017-ok] few, if any, have explored improving forecast accuracy using a mixed effects framework. Mixed effects models are useful when non-independence is observed in the data, which often occurs when multiple observations are recorded from the same subject. If buildings behave in a similar manner a well-specified mixed model may produce more accurate forecasts than individual models. Furthermore, a mixed-model framework allows us to quantify differences between buildings which would not otherwise be possible when using a "building-specific" modelling approach. A mixed effects approach opens the door to scenario analyses by allowing us to estimate how demand might change under different equipment or usage scenarios.

@@

Several linear and mixed effects models were tested to determine which produced the most accurate forecasts conditional on selected features. Here we describe the various benchmarking and mixed effects models and their formulations. All analysis was produced using the R statistical programming language [@R-base]. Mixed effects models were fit using the `lme4` package for mixed effects models [@R-lme4]. All code used to run this analysis has been made available in the [https://github.com/camroach87/1901-nlmets](https://github.com/camroach87/1901-nlmets) GitHub repo.
